### 2025-09-19 YEAR2/AER210_LectureNotes/main.tex
sept19

\subsection{Triple Integrals}
The idea of a triple integral (in fact n-tuple integral) could be extend from the idea of double integral, similar to 
Definition \ref{def:double_integral}. We can define a triple integral as follows:
\begin{definition}
    Consider a function $f(x,y,z)$ that is continuous a 
    3-D reigion with volumne $V$. We can partition the reigion into $n$ subreigons with volume $\Delta V_i$. Then, we can define the triple integral of $f$ over the reigion $V$ as follows:
    \begin{equation}
        \iiint_V f(x,y,z) \, dV = \lim_{|P| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i
    \end{equation}
    where $(x_i^*, y_i^*, z_i^*)$ is a point in the $i$-th subregion, and $|P|$ is the maximum diameter of the subregions. This limit exists and equals to the triple integral if $f$ is continuous on $V$, since the following squeeze theorem holds, similar to the double integral case:
    $$       
     \sum_{i=1}^n m_i \Delta V_i \le \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i \le \sum_{i=1}^n M_i \Delta V_i
    $$
    where $m$ and $M$ are the minimum and maximum values of $f$ on $V$, respectively.

@@

\subsection{Triple Integrals in Rectangular Coordinates}
\begin{definition}[Triple Integral in Rectangular Region]
    Consider a function $f(x,y,z)$ that is continuous over a rectangular box $B$ defined by the inequalities:
    $$
        a \le x \le b, \quad c \le y \le d, \quad r \le z \le s
    $$
    Then, the triple integral of $f$ over the box $B$ can be computed as an iterated integral in any order of integration:
    \begin{equation}
        \iiint_B f(x,y,z) \, dV = \int_a^b \int_c^d \int_r^s f(x,y,z) \, dz \, dy \, dx
    \end{equation}
    or any other permutation of the order of integration.

@@

    First, we select the $z = 4 - y^2$ surface as our base reigion $R$. Then, we consider how the solid change in the $x$ direction over $R$:
    $$
        x + z = 4 \implies 0 \le x \le 4 - z
    $$
    So we can set up:
    \begin{align*}
        V &= \iint_R \int_0^{4 - z} 1 \, dx \, dA \\
        &= \iint_R (4 - z) \, dA
    \end{align*}
    Now, we consider how the projection $R$ change in the $z$ direction over $y$:
    $$        z = 4 - y^2 \implies 0 \le z \le 4 - y^2, \quad -2 \le y \le 2
    $$
    So we can set up:
    \begin{align*}
        V &= \int_{-2}^2 \int_0^{4 - y^2} (4 - z) \, dz \, dy \\
        &= \int_{-2}^2 \left[ 4z - \frac{z^2}{2} \right]_{z=0}^{z=4 - y^2} \, dy \\
        &= \int_{-2}^2 \left( 8 - \frac{y^2}{2} \right) \, dy \\
        &= \frac{128}{5}
    \end{align*}
\end{example}
\subsection{Triple Integrals in Spherical Coordinates}
\subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
\chapter{Fluid Mechanics}
\end{document}

>>>>

\subsection{Triple Integrals in Rectangular Coordinates}
The idea of a triple integral (in fact n-tuple integral) could be extend from the idea of double integral, similar to 
Definition \ref{def:double_integral}. We can define a triple integral as follows:
\begin{definition}
    Consider a function $f(x,y,z)$ that is continuous a 
    3-D reigion with volumne $V$. We can partition the reigion into $n$ subreigons with volume $\Delta V_i$. Then, we can define the triple integral of $f$ over the reigion $V$ as follows:
    \begin{equation}
        \iiint_V f(x,y,z) \, dV = \lim_{|P| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i
    \end{equation}
    where $(x_i^*, y_i^*, z_i^*)$ is a point in the $i$-th subregion, and $|P|$ is the maximum diameter of the subregions. This limit exists and equals to the triple integral if $f$ is continuous on $V$, since the following squeeze theorem holds, similar to the double integral case:
    $$       
     \sum_{i=1}^n m_i \Delta V_i \le \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i \le \sum_{i=1}^n M_i \Delta V_i
    $$
    where $m$ and $M$ are the minimum and maximum values of $f$ on $V$, respectively.

@@

\begin{definition}[Triple Integral in Rectangular Region]
    Consider a function $f(x,y,z)$ that is continuous over a rectangular box $B$ defined by the inequalities:
    $$
        a \le x \le b, \quad c \le y \le d, \quad r \le z \le s
    $$
    Then, the triple integral of $f$ over the box $B$ can be computed as an iterated integral in any order of integration:
    \begin{equation}
        \iiint_B f(x,y,z) \, dV = \int_a^b \int_c^d \int_r^s f(x,y,z) \, dz \, dy \, dx
    \end{equation}
    or any other permutation of the order of integration.

@@

    First, we select the $z = 4 - y^2$ surface as our base reigion $R$. Then, we consider how the solid change in the $x$ direction over $R$:
    $$
        x + z = 4 \implies 0 \le x \le 4 - z
    $$
    So we can set up:
    \begin{align*}
        V &= \iint_R \int_0^{4 - z} 1 \, dx \, dA \\
        &= \iint_R (4 - z) \, dA
    \end{align*}
    Now, we consider how the projection $R$ change in the $z$ direction over $y$:
    $$        z = 4 - y^2 \implies 0 \le z \le 4 - y^2, \quad -2 \le y \le 2
    $$
    So we can set up:
    \begin{align*}
        V &= \int_{-2}^2 \int_0^{4 - y^2} (4 - z) \, dz \, dy \\
        &= \int_{-2}^2 \left[ 4z - \frac{z^2}{2} \right]_{z=0}^{z=4 - y^2} \, dy \\
        &= \int_{-2}^2 \left( 8 - \frac{y^2}{2} \right) \, dy \\
        &= \frac{128}{5}
    \end{align*}
\end{example}

\begin{example}
    \textit{Change the order of Integration in the following triple integral such that the integration is performed w.r.t. $x$, then $y$, then $z$:}
    $$
        \int_{-1}^1 \int_{x^2}^1 \int_0^{1 - y} f(x,y,z) \, dz \, dy \, dx
    $$
    The region is limited in the $z$ direction by $0 \le z \le 1 - y$:
    % use plot 3d to draw z = 1 - y and z = 0
    \begin{figure}[h!]
        \centering
        \begin{tikzpicture}
            \begin{axis}[
                    axis lines = middle,
                    xlabel = $x$,
                    ylabel = $y$,
                    zlabel = $z$,
                    domain=-1:1,
                    y domain=0:1,
                    samples=20,
                    samples y=10,
                    colormap/cool,
                    view={120}{30},
                    width=10cm,
                    height=8cm,
                ]
                % Draw the surface z = 1 - y
                \addplot3[surf, opacity=0.5] {1 - y};
                % Draw the plane z = 0
                \addplot3[surf, opacity=0.5] {0};
                % Draw the curve y = x^2
                \addplot3[domain=-1:1, samples=100, thick, blue] ({x}, {x^2}, {0});
                % Draw the vertical lines at x = -1 and x = 1
                \addplot3[domain=0:1, samples=100, thick, red] ({-1}, {x}, {0});
                \addplot3[domain=0:1, samples=100, thick, red] ({1}, {x}, {0});
            \end{axis}
        \end{tikzpicture}
        \caption{The region bounded by $z = 1 - y$, $z = 0$, $y = x^2$, $x = -1$, and $x = 1$.}
    \end{figure}

    We can also visualize the $R$ that is the reigion projected in the $xy$-plane as follows

    \begin{figure}[h!]
        \centering
        \begin{tikzpicture}[scale=2]
            % Draw axes
            \draw[->] (-1.5,0) -- (1.5,0) node[right] {$x$};
            \draw[->] (0,-0.5) -- (0,1.5) node[above] {$y$};

            % Draw the curve y = x^2
            \draw[domain=-1.2:1.2,smooth,variable=\x,blue] plot ({\x},{\x*\x}) node[right] {$y = x^2$};
            % Draw the line y = 1
            \draw[blue] (-1,1) -- (1,1) node[right] {$y = 1$};

            %vertical labels for x = pm1
            \draw[dashed] (-1,0) -- (-1,1);
            \draw[dashed] (1,0) -- (1,1);
            \node at (-1,-0.1) {$-1$};
            \node at (1,-0.1) {$1$};

            % Shade the region
            \fill[blue!20,opacity=0.5] plot[domain=-1:1,smooth,variable=\x] ({\x},{\x*\x}) -- (1,1) -- (-1,1) -- cycle;
            % Add labels
            \node at (0.5,0.5) {$R$};
        \end{tikzpicture}
        \caption{The projection of the region onto the $xy$-plane.}

    \end{figure}

    So, we can consider the projection of the triangle on the $zy$-plane as our base region $R'$. Then, we consider how the region changes in the $x$ direction over $R$:
    $$
        x^2 = y \implies -\sqrt{y} \le x \le \sqrt{y}
    $$
    So we can set up:
    \begin{align*}
        \iiint_V f(x,y,z) \, dV &= \iint_{R'} \int_{-\sqrt{y}}^{\sqrt{y}} f(x,y,z) \, dx \, dA \\
        &= \iint_{R'} \int_{-\sqrt{y}}^{\sqrt{y}} f(x,y,z) \, dx \, dy \, dz
    \end{align*}
    Now, we consider how the projection $R'$ change in the $y$ direction over $z$:
    $$        z = 1 - y \implies 0 \le y \le 1 - z, \quad 0 \le z \le 1
    $$
    So we can set up:
    \begin{align*}
        \iiint_V f(x,y,z) \, dV &= \int_0^1 \int_0^{1 - z} \int_{-\sqrt{y}}^{\sqrt{y}} f(x,y,z) \, dx \, dy \, dz
    \end{align*}
\end{example}
\subsubsection{Applications of Triple Integrals}
\paragraph{Physical Applications} Similar to double integrals, triple integrals can be used to compute physical quantities such as mass, center of mass, and moment of inertia for three-dimensional objects with variable density. 
\begin{definition}[Mass of a Solid]
    Consider a solid $V$ with a variable density function $\rho(x,y,z)$ that is continuous over the solid. The mass $M$ of the solid can be computed using the following triple integral:
    \begin{equation}
        M = \iiint_V \rho(x,y,z) \, dV
    \end{equation}

\begin{definition}[Center of Mass of a Solid]
    Consider a solid $V$ with a variable density function $\rho(x,y,z)$ that is continuous over the solid. The coordinates of the center of mass $(\bar{x}, \bar{y}, \bar{z})$ of the solid can be computed using the following formulas:
    \begin{equation}
        \bar{x} = \frac{1}{M} \iiint_V x \rho(x,y,z) \, dV, \quad \bar{y} = \frac{1}{M} \iiint_V y \rho(x,y,z) \, dV, \quad \bar{z} = \frac{1}{M} \iiint_V z \rho(x,y,z) \, dV
    \end{equation}
    where $M$ is the mass of the solid.

\begin{definition}[Moment of Inertia of a Solid]
    Consider a solid $V$ with a variable density function $\rho(x,y,z)$ that is continuous over the solid. The moment of inertia of the solid about the $x$-axis ($I_x$), $y$-axis ($I_y$), and $z$-axis ($I_z$) can be computed using the following formulas:
    \begin{equation}
        I_x = \iiint_V (y^2 + z^2) \rho(x,y,z) \, dV, \quad I_y = \iiint_V (x^2 + z^2) \rho(x,y,z) \, dV, \quad I_z = \iiint_V (x^2 + y^2) \rho(x,y,z) \, dV
    \end{equation}
    where $y^2 + z^2$, $x^2 + z^2$, and $x^2 + y^2$ are the squares of the distances from the respective axes of rotation, and $\rho(x,y,z) \, dV$ represents the mass element of the solid at the point $(x,y,z)$.

    Also for the moment of inertia about the origin ($I_o$):
    \begin{equation}
        I_o = \iiint_V (x^2 + y^2 + z^2) \rho(x,y,z) \, dV
    \end{equation}
    In general, for an axis defined by a line $ax + by + cz + d = 0$, the moment of inertia about that axis ($I_l$) can be found using the following formula:
    \begin{equation}
        I_l = \iiint_V \left( \frac{ax + by + cz + d}{\sqrt{a^2 + b^2 + c^2}} \right)^2 \rho(x,y,z) \, dV
    \end{equation}
\end{definition}

\begin{example}
    \textit{Find the center of mass of a solid of constant density $\rho_0$ that is bounded by the parabolic cylinder $x = y^2$ and the planes $z=0$, $x=z$ and $x=1$.}
    We visualize the solid as follows:
    \begin{figure}[h!]
        \centering
        \begin{tikzpicture}
            \begin{axis}[
                    axis lines = middle,
                    xlabel = $x$,
                    ylabel = $y$,
                    zlabel = $z$,
                    domain=0:1.3,
                    y domain=-1:1,
                    samples=20,
                    samples y=10,
                    colormap/cool,
                    view={120}{30},
                    width=10cm,
                    height=8cm,
                ]
                % Draw the surface z = x
                \addplot3[surf, opacity=0.5] {x};
                % Draw the plane z = 0
                \addplot3[surf, opacity=0.5] {0};

                % Draw the surface cylinder x = y^2 on the plane z = 0, and it extends to z = x
                % Parabolic cylinder x = y^2 (y in [-1,1], z in [0,1])
                \addplot3[surf, opacity=0.45, shader=flat, domain=-1:1, y domain=0:1] ({x^2}, {x}, {y});
                % bottom z = 0
                \addplot3[surf, opacity=0.25, shader=flat, domain=0:1, y domain=-1:1] ({x}, {y}, {0});
                % slanted plane x = z (i.e. z = x)
                \addplot3[surf, opacity=0.35, shader=flat, domain=0:1, y domain=-1:1] ({x}, {y}, {x});
                % vertical plane x = 1 (closes the solid)
                \addplot3[surf, opacity=0.30, shader=flat, domain=-1:1, y domain=0:1] ({1}, {x}, {y});
                % fill the surface such that
            \end{axis}
        \end{tikzpicture}
        \caption{The solid bounded by the surfaces $x = y^2$, $z = 0$, $x = z$, and $x = 1$.}
    \end{figure}
    We can find the mass of the solid:
    \begin{align*}
        M &= \iiint_V \rho_0 \, dV = \rho_0 \iiint_V 1 \, dV \\
        &= \rho_0 \int_{-1}^1 \int_{y^2}^1 \int_{0}^x 1 \, dz \, dx \, dy \\
        &= \rho_0 \cdot \frac{4}{5}
    \end{align*}

    Next, we can find the coordinates of the center of mass:
    \begin{align*}
        \bar{x} &= \frac{1}{M} \iiint_V x \rho_0 \, dV = \frac{\rho_0}{M} \iiint_V x \, dV = 5/7 \\
        \bar{y} &= \frac{1}{M} \iiint_V y \rho_0 \, dV = \frac{\rho_0}{M} \iiint_V y \, dV = 0 \text{ (by symmetry)} \\
        \bar{z} &= \frac{1}{M} \iiint_V z \rho_0 \, dV = \frac{\rho_0}{M} \iiint_V z \, dV = 5/14 \\
    \end{align*}    
\end{example}

\begin{example}
    \textit{Find the moment of inertia of a cyliner with radius $a$ and height $h$ about its central axis, assuming the cylinder has a constant density $\rho_0$.}
    We consider the base region $R$ as the circular base of the cylinder in the $xy$-plane. Then, we consider how the cylinder change in the $z$ direction over $R$:
    $$
        0 \le z \le h
    $$
    So we can set up:
    \begin{align*}
        I_z &= \iiint_V (x^2 + y^2) \rho_0 \, dV = \rho_0 \iint_R \int_0^h (x^2 + y^2) \, dz \, dA \\
        &= \rho_0 h \iint_R (x^2 + y^2) \, dA
    \end{align*}

    Then, we can express the base region $R$ in polar coordinates:
    $$        
    x^2 + y^2 = r^2 \implies 0 \le r \le a, \quad 0 \le \theta \le 2\pi
    $$
    So we can set up:
    \begin{align*}
        I_z &= \rho_0 h \int_0^{2\pi} \int_0^a r^2 \cdot r \, dr \, d\theta \\
        &= \rho_0 h \int_0^{2\pi} \left[ \frac{r^4}{4} \right]_{r=0}^{r=a} \, d\theta = \frac{\pi \rho_0 h a^4}{2}
    \end{align*}
    This also leads to the well-known formula for the moment of inertia of a solid cylinder about its central axis:
    $$
        I_z = \frac{1}{2} M a^2
    $$
    where $M = \pi a^2 h \rho_0$ is the mass of the cylinder. 
\end{example}
\subsection{Triple Integrals in Cylindrical Coordinates}
\begin{definition}[Cylindrical Coordinates]
    The cylindrical coordinates of a point $P$ in 3-D space are given by the ordered triple $(r, \theta, z)$, where:
    \begin{itemize}
        \item $r$ is the distance from the $z$-axis to the projection of $P$ onto the $xy$-plane,
        \item $\theta$ is the angle between the positive $x$-axis and the line segment from the origin to the projection of $P$ onto the $xy$-plane,
        \item $z$ is the same as in rectangular coordinates, representing the height of point $P$ above the $xy$-plane.
    \end{itemize}
    The relationships between cylindrical coordinates and rectangular coordinates are given by:
    \begin{equation}
        x = r \cos \theta, \quad y = r \sin \theta, \quad z = z
    \end{equation}
    and conversely:
    \begin{equation}
        r = \sqrt{x^2 + y^2}, \quad \theta = \tan^{-1}\left(\frac{y}{x}\right), \quad z = z
    \end{equation}
\end{definition}

\begin{example}

\end{example}
\subsection{Triple Integrals in Spherical Coordinates}
\subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
\chapter{Fluid Mechanics}
\end{document}

### 2025-09-19 YEAR2/CHE260_LectureNotes/main.tex
sept19

\begin{example}
    Consider a piston-cylinder device. If we put a weight on top, the gas is compressed, and the piston moves down. This boundry work process would change enthalpy, despite not changing the temperature. This is becuase:
    $$
        dh = du + PdV + VdP = VdP \quad (du = 0, dV = 0)
    $$
\end{example}
\chapter{Heat Transfer}

>>>>

\begin{example}
    Consider a piston-cylinder device. If we put a weight on top, the gas is compressed, and the piston moves down. This boundry work process would change enthalpy, despite not changing the temperature. This is becuase:
    $$
        dh = du + PdV + VdP = VdP \quad (du = 0, dV = 0)
    $$
\end{example}

\begin{example}[Control Volume in a Turbine]
    The length of a fluid inside of control volume given time is:
    $$
        dx = V dt
    $$
    The volume is given by:
    $$
        dV = A dx = A V dt
    $$
    The mass of the fluid is:
    $$
        \delta m = \rho dV = \rho A V dt
    $$
    We can infer that if work is done, then the pressure is non constant. And that:
    $$
        \dot{m} = \rho A V = \text{constant}
    $$
    So at the inlet we have high $\rho$ and high $P$, and at the outlet we have low $\rho$ and low $P$. Such that $V$ remains relatively constant. We also devucebce specific volume:
    \begin{equation}
        v = \frac{1}{\rho}
    \end{equation}
    So we have:
    \begin{equation}
        \dot{m} = \frac{A V}{v} = \text{constant}
    \end{equation}
    At steady state:
    \begin{equation}
        \dot{m}_{in} = \dot{m}_{out}
    \end{equation}
    For and ideal gas, we have:
    \begin{equation}
        Pv = RT
    \end{equation}
\end{example}
\begin{definition}[Fluid Energy per Unit Mass]
    The total fluid energy per unit mass is given by:
    \begin{equation}
        e = h + \frac{V^2}{2} + gz
    \end{equation}
    where $h$ is the specific enthalpy, $\frac{V^2}{2}$ is the kinetic energy per unit mass, and $gz$ is the potential energy per unit mass.

    Now, we can write the energy balance at the steady state, for a steady flow device:
    \begin{equation}
        q + w = (h_2 - h_1) + \left(\frac{V_2^2}{2} - \frac{V_1^2}{2}\right) + g(z_2 - z_1)
    \end{equation}

    For turbine, assume $q = 0$, and the KE and PE are small, we have:
    \begin{equation}
        w = h_1 - h_2, \quad W = m(h_1 - h_2)
    \end{equation}
    For compressor, we have $w_{shaft} = h_2 - h_1 > 0$.

\begin{example}
    Consider a compression turbine that has 800$^\circ$C and 600 kPa at the inlet, and 300$^\circ$C and 100 kPa at the outlet. Assuming ideal gas behavior, and $c_{p, \text{air}} = 1.005$ kJ/kg K. A 10kg flow rate would have:
    $$
        w = h_1 - h_2 
    $$
    We can compute the enthalpy change:
    $$
        \Delta h = \int_{T_1}^{T_2} c_p(T) dT \approx c_p (T_2 - T_1) = 1.005 \times (573 - 1073) = -502.5\, \text{kJ/kg}
    $$
    So the power output is:
    $$
        W = m \Delta h = 10 \times -502.5 = -5.025\, \text{MW}
    $$
\end{example}

\begin{example}[Pump at Constant Temperature]
    Consider a pump (a up pumping pump) that has 20$^\circ$C and 100 kPa at the inlet, and 20$^\circ$C and 600 kPa at the outlet. Assuming incompressible liquid water, and $v = 0.001\, \text{m}^3/\text{kg}$. A 10kg flow rate would have:
    $$
        w_{shaft} = h_2 - h_1 + g(z_2 - z_1)
    $$
    We can compute the enthalpy change:
    $$
        \Delta h = v (P_2 - P_1) 
    $$
    So the power with considerion of the change in potential energy is:
    $$
        W = m \Delta h + mg(z_2 - z_1) = 10 \times 0.001 \times (600 - 100) + 10 \times 9.81 \times (10 - 0) = 98.5\, \text{kW}
    $$
\end{example}
\chapter{Heat Transfer}

### 2025-09-19 YEAR2/ECE253_LectureNotes/main.tex
sept19

\begin{definition}[Continuous Assignment]
    A continuous assignment is used to model combinational logic in Verilog. It is defined using the \verb|assign| keyword, followed by the output signal, the assignment operator `=`, and the logic expression. Continuous assignments are evaluated whenever any of the input signals change. That is, the output is considered instantaneously updated when the input changes (ignoring propagation delay).
\end{definition}

\begin{example}[2-1 Multiplexers (Mux)]
    \textit{Design a circuit that controls a light $f$ based on either two switches $x$ and $y$. The switch that control the light is determined by a control signal $s$. If $s = 0$, the light is controlled by switch $x$. If $s = 1$, the light is controlled by switch $y$.}
    \\
    We have the following truth table:

@@

\paragraph{The diagram of a Mux} is a trapezoid shown below:
\begin{center}
    \begin{tikzpicture}[scale=1]
        % Input nodes
        \node (x) at (-2,0.5) {$x$};
        \node (y) at (-2,-0.5) {$y$};
        \node (s) at (-0.5,1.5) {$s$}; 
        % Wires to gate
        \draw (x.east) -- (-1,0.5);
        \draw (y.east) -- (-1,-0.5);
        \draw (s.south) -- (-0.5,0.75);
        % Gate body: trapezoid
        \draw (-1,1) -- (0,0.5) -- (0,-0.5) -- (-1,-1) -- cycle;
        % Output
        \node (f) at (1,0) {$f$};

@@

\paragraph{4-1 Multiplexer (Mux)} A 4-1 Mux has 4 data inputs ($d_0, d_1, d_2, d_3$). The select signal would have two bit ($s_0, s_1$) to select one of the four data inputs to be outputted. In general, this is call a multibit signal, a.k.a a bus.
\begin{definition}

>>>>

\begin{definition}[Continuous Assignment]
    A continuous assignment is used to model combinational logic in Verilog. It is defined using the \verb|assign| keyword, followed by the output signal, the assignment operator `=`, and the logic expression. Continuous assignments are evaluated whenever any of the input signals change. That is, the output is considered instantaneously updated when the input changes (ignoring propagation delay).
\end{definition}
\subsection{Multiplexers (Mux)}
\begin{example}[2-1 Multiplexers (Mux)]
    \textit{Design a circuit that controls a light $f$ based on either two switches $x$ and $y$. The switch that control the light is determined by a control signal $s$. If $s = 0$, the light is controlled by switch $x$. If $s = 1$, the light is controlled by switch $y$.}
    \\
    We have the following truth table:

@@

\paragraph{The Diagram of a Mux} is a trapezoid shown below:
\begin{center}
    \begin{tikzpicture}[scale=1]
        % Input nodes
        \node (x) at (-2,0.5) {$x$};
        \node (y) at (-2,-0.5) {$y$};
        \node (s) at (-0.5,1.5) {$s$}; 
        % Wires to gate
        \draw (x.east) -- (-1,0.5);
        \draw (y.east) -- (-1,-0.5);
        \draw (s.south) -- (-0.5,0.75);
        % Gate body: trapezoid
        \draw (-1,1) -- (0,0.5) -- (0,-0.5) -- (-1,-1) -- cycle;
        % Output
        \node (f) at (1,0) {$f$};

@@

\paragraph{4-1 Multiplexer (Mux)} A 4-1 Mux has 4 data inputs ($d_0, d_1, d_2, d_3$). The select signal would have two bit ($s_0, s_1$) to select one of the four data inputs to be outputted. In general, this is call a multibit signal, a.k.a a bus.
\begin{definition}[Bus]
    A bus is a bundle of signals. It is used to transfer data between different components of a digital system.
\end{definition}

\begin{example}[Verilog for a 2bit 2-1 Mux]
    The verilog implementation for a 2bit 2-1 Mux is as follows:
    \begin{verbatim}
    module mux2to1_2bit(input logic [1:0] x, y, // 2-bit inputs
                        input logic s,         // select signal
                        output logic [1:0] f); // 2-bit output
        assign f[0] = (~s & x[0]) | (s & y[0]); // LSB
        assign f[1] = (~s & x[1]) | (s & y[1]); // MSB
        // Alternatively, we can use the following single line:
        // assign f = s ? y : x; // If s=1, f=y; else f=x
    endmodule
    \end{verbatim}

    As we can see, we can initialize the datatype of a bus using the following syntax:
    \begin{verbatim}
    logic [n-1:0] bus_name; // n-bit bus
    \end{verbatim}

\subsection{Adders}
\begin{definition}[Half Adder]
    A half adder is a combinational logic circuit that performs the addition of two single-bit binary numbers. It has two inputs, typically denoted as $A$ and $B$, and two outputs: the sum ($S$) and the carry ($C$). The sum output represents the least significant bit of the addition, while the carry output represents any overflow that occurs when both inputs are 1. We start with the following truth table:
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c|c|c|}
            \hline
            $A$ & $B$ & $S_1$ & $S_2$ \\
            \hline
            0 & 0 & 0 & 0 \\
            0 & 1 & 1 & 0 \\    
            1 & 0 & 1 & 0 \\
            1 & 1 & 0 & 1 \\
            \hline
        \end{tabular}
        \caption{Truth Table for Half Adder}
        \label{tab:half_adder_truth_table}
    \end{table}
    From the truth table, we can derive the following expressions using canonical SOP form:
    \begin{align*}
        S_0 &= m_1 + m_2 = \overline{A}B + A\overline{B} = A \oplus B \\
        S_1 &= m_3 = AB
    \end{align*}
    The Verilog implementation is as follows:
    \begin{verbatim}
    module half_adder(input logic A, B,
                      output logic [1:0] S);
        assign s[1] = a & b;
        assign s[0] = a ^ b;
    endmodule
    \end{verbatim}
\end{definition}

\begin{definition}[Full Adder]
    A full adder is a combinational logic circuit that performs the addition of three single-bit binary numbers: two significant bits and a carry-in bit. It has three inputs, we denote the signifcant bit as $A_i$ and $B_i$, and the carry-in bit as $C_i$, and two outputs: the sum ($S_i$) and the carry-out ($C_{i + 1}$). The sum output represents the least significant bit of the addition, while the carry-out output represents any overflow that occurs when the sum exceeds the value that can be represented by a single bit. We start with the following truth table:
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c|c|c|c|}
            \hline
            $A_i$ & $B_i$ & $C_i$ & $S_i$ & $C_{i+1}$ \\
            \hline
            0 & 0 & 0 & 0 & 0 \\
            0 & 0 & 1 & 1 & 0 \\
            0 & 1 & 0 & 1 & 0 \\
            0 & 1 & 1 & 0 & 1 \\
            1 & 0 & 0 & 1 & 0 \\
            1 & 0 & 1 & 0 & 1 \\
            1 & 1 & 0 & 0 & 1 \\
            1 & 1 & 1 & 1 & 1 \\
            \hline
        \end{tabular}
        \caption{Truth Table for Full Adder}
        \label{tab:full_adder_truth_table}
    \end{table}
    From the truth table, we can derive the following expressions using canonical SOP form:
    \begin{align*}
        S_i &= m_1 + m_2 + m_4 + m_7 \\
            &= \overline{A_i}\overline{B_i}C_i + \overline{A_i}B_i\overline{C_i} + A_i\overline{B_i}\overline{C_i} + A_iB_iC_i \\
            &= A_i \oplus B_i \oplus C_i \\
        C_{i+1} &= m_3 + m_5 + m_6 + m_7 \\
                &= \overline{A_i}B_iC_i + A_i\overline{B_i}C_i + A_iB_i\overline{C_i} + A_iB_iC_i \\
    \intertext{Using the Distributive Law and A = A + A}
                &= B_iC_i(\overline{A_i} + A_i) + A_iB_i(\overline{C_i} + C_i) + A_iC_i(\overline{B_i} + B_i) \\
                &= A_iB_i + B_iC_i + A_iC_i
    \end{align*}
    The circuit diagram of a full adder is shown below:
    \begin{figure}[h!]
        \begin{circuitikz}
            \draw
            % gates 
            (0,0) node[xor port] (X1) {}
            (2,0) node[xor port] (X2) {}
            (1,-2) node[and port] (A1) {}
            (3,-2) node[and port] (A2) {}
            (2,-4) node[and port] (A3) {}
            (4,-4) node[or port] (O1) {} 
            (6,-4) node[or port] (O2) {};

            % connections
            \draw (X1.out) -- (X2.in 1);
            \draw (X2.out) -- ++(1,0) node[right] {$S_i$};

            % inputs
            \draw (X1.in 1) -- ++(-1,0) node[left] {$A_i$};
            \draw (X1.in 2) -- ++(-1,0) node[left] {$B_i$};
            \draw (X2.in 2) -- ++(-0.3,0) node[left] {$C_i$};

            % AND gate connections
            \draw (X1.in 1) |- (A1.in 1);
            \draw (X1.in 2) |- (A1.in 2);
            \draw (X1.in 2) |- (A2.in 1);
            \draw (X2.in 2) |- (A2.in 2);
            \draw (X1.in 1) |- (A3.in 1);
            \draw (X2.in 2) |- (A3.in 2);

            % OR gate connections
            \draw (A1.out) |- (O1.in 1);
            \draw (A2.out) |- (O1.in 2);
            \draw (A3.out) |- (O2.in 1);
            \draw (O1.out) |- (O2.in 2);
            \draw (O2.out) -- ++(1,0) node[right] {$C_{i+1}$}
        ;

        \end{circuitikz}
        \caption{Circuit Diagram of a Full Adder}
        \label{fig:full_adder_circuit}
        \centering 
    \end{figure}

    The Verilog implementation is as follows:
    \begin{verbatim}
    module full_adder(input logic A_i, B_i, C_i,
                      output logic S_i, C_i_plus_1)
        assign S_i = A_i ^ B_i ^ C_i;
        assign C_i_plus_1 = (A_i & B_i) | (B_i & C_i) | (A_i & C_i);
    endmodule
    \end{verbatim}
\end{definition}

\begin{definition}[Hierarchical Verilog Module]
    Hierarchical Verilog is a verilog module that instantiates other modules within it. This allows for the creation of complex designs by combining simpler modules.
\end{definition}
\begin{example}[3-bit Ripple Carry Adder]
    A 3-bit ripple carry adder can be implemented using three instances of the full adder module. The Verilog implementation is as follows:
    \begin{verbatim}
    module adder3(input logic [2:0] A, B,
                  input logic C_in,
                  output logic [2:0] S,
                  output logic C_out);
        logic C_1, C_2; // Internal carry signals

        // Instantiate full adders
        full_adder FA0 (A[0], B[0], C_in, S[0], C_1);
        full_adder FA1 (A[1], B[1], C_1, S[1], C_2);
        full_adder FA2 (A[2], B[2], C_2, S[2], C_out);
    endmodule
    \end{verbatim}
\end{example}
